\name: A flexible and data-driven $K$-process model for nucleosynthesis

### 2023-06-17 Paper/main.tex
various content and references

\documentclass[modern]{aastex631}
\usepackage{amsmath}
\usepackage{soul}

% Abundance and stellar macros
\newcommand{\mgfe}[0]{[{\rm Mg/Fe}]}
\newcommand{\afe}[0]{[\alpha/{\rm Fe}]}
\newcommand{\femg}{[{\rm Fe}/{\rm Mg}]} 
\newcommand{\xmg}{[{\rm X}/{\rm Mg}]} 
\newcommand{\mgh}{[{\rm Mg}/{\rm H}]}
\newcommand{\feh}[0]{[{\rm Fe/H}]}
\newcommand{\xfe}{[{\rm X}/{\rm Fe}]}
\newcommand{\xh}{[{\rm X}/{\rm H}]}

\newcommand{\logg}{\log(g)}
\newcommand{\teff}{T_{\rm eff}}
\newcommand{\kpc}{\rm \; kpc}
\newcommand{\kel}{\rm \; K}
\newcommand{\msun}{M_{\odot}}
\newcommand{\rgc}{R_{\rm GC}}
\newcommand{\msunvice}{M_{\odot}/M_{\odot \rm formed}}
\newcommand{\MZAMS}{M_{\rm ZAMS}}

\newcommand{\qcc}{q_{{\rm CC,}j}^{Z}}
\newcommand{\qccFe}{q_{{\rm CC, Fe}}^{Z}}
\newcommand{\dqccFe}{dq_{{\rm CC, Fe}}^{Z}/dZ}
\newcommand{\qIa}{q_{{\rm Ia,}j}^{Z}}
\newcommand{\qIaFe}{q_{{\rm Ia,Fe}}^{Z}}
\newcommand{\Acc}{A^{\rm CC}_{i}}
\newcommand{\AIa}{A^{\rm Ia}_{i}}
\newcommand{\fcc}{f_{{\rm CC}, ij}}

@@

After hydrogen, helium, lithium, and beryllium, all other naturally occurring elements are made in the stars, and the collisions of stars.
Stellar surface abundances---the abundances measured by taking a spectrum of a stellar photosphere---are thought to deliver a relatively unprocessed record of the element abundances in the gas from which the star formed \citep[though see e.g.,][]{pinsonneault2001, oh2018, souto2019, vincenzo2021b}.
These birth abundances were set by a combination of nucleosynthetic processes involved in making heavy atomic nuclei, and astrophysical processes involved in delivering atoms from stellar interiors to star-formation sites \citep[e.g.,][]{johnsonja2020}.
Thus nuclear physics and a wide swath of astrophysics are critically intertwined in our understanding of stellar surface abundances, motivating theoretical, experimental, and observational work.

@@

\paragraph{3. Non-negativity}
All process vector components for all elements are non-negative and all process amplitudes are non-negative.
This assumption implies that the elements considered here are only produced, and not ever destroyed, by the $K$ processes (relative to hydrogen).
This makes the model similar to a non-negative matrix factorization (\citealt{nnmf} \ejg{Hogg add citation}). In \name{}, this assumption is enforced by requiring that the process vector components and amplitudes are always greater than or equal to zero, such that
\begin{equation} \label{eq:constraint_zero}
    q_{k,j}^Z \geq 0 \; \forall \; Z,k,j \quad \text{and} \quad A_i^k \geq 0 \; \forall \; k,i ~.
\end{equation}

@@

Because we don't want to be too drawn or influenced by outlier points, we don't use the observed errors $\sigma_{{\rm obs},ij}$ in the likelihood, but instead we soften them in the spirit of iteratively reweighted least squares (HOGG CITE):
\begin{equation}\label{eq:inflate_ivar}
    \frac{1}{\sigma_{ij}^2} = \frac{Q^2/\sigma_{{\rm obs},ij}^2}{Q^2 + (\xh_{ij} - m_{ij})^2 /\sigma_{{\rm obs},ij}^2 },    
\end{equation}
where $Q$ is a softening parameter, which is set to $Q=5$.

@@

\bibliography{sample631}{}
\bibliographystyle{aasjournal}

>>>>

% This document is part of the KPM project
% Copyright 2023 the authors.
% ------------------------------------------------------------------

% to-do items
% -----------
% - Look at all latex warnings on the build

\documentclass[modern]{aastex631}
\usepackage{amsmath}
\usepackage{microtype}
\usepackage{soul}

% Abundance and stellar macros
\newcommand{\mgfe}[0]{[{\rm Mg/Fe}]}
\newcommand{\afe}[0]{[\alpha/{\rm Fe}]}
\newcommand{\femg}{[{\rm Fe}/{\rm Mg}]} 
\newcommand{\xmg}{[{\rm X}/{\rm Mg}]} 
\newcommand{\mgh}{[{\rm Mg}/{\rm H}]}
\newcommand{\feh}[0]{[{\rm Fe/H}]}
\newcommand{\xfe}{[{\rm X}/{\rm Fe}]}
\newcommand{\xh}{[{\rm X}/{\rm H}]}
\newcommand{\logg}{\log(g)}
\newcommand{\teff}{T_{\rm eff}}
\newcommand{\kpc}{\rm \; kpc}
\newcommand{\kel}{\rm \; K}
\newcommand{\msun}{M_{\odot}}
\newcommand{\rgc}{R_{\rm GC}}
\newcommand{\msunvice}{M_{\odot}/M_{\odot \rm formed}}
\newcommand{\MZAMS}{M_{\rm ZAMS}}
\newcommand{\qcc}{q_{{\rm CC,}j}^{Z}}
\newcommand{\qccFe}{q_{{\rm CC, Fe}}^{Z}}
\newcommand{\dqccFe}{dq_{{\rm CC, Fe}}^{Z}/dZ}
\newcommand{\qIa}{q_{{\rm Ia,}j}^{Z}}
\newcommand{\qIaFe}{q_{{\rm Ia,Fe}}^{Z}}
\newcommand{\Acc}{A^{\rm CC}_{i}}
\newcommand{\AIa}{A^{\rm Ia}_{i}}
\newcommand{\fcc}{f_{{\rm CC}, ij}}

@@

After hydrogen, helium, lithium, and beryllium, all other naturally occurring elements are made in stars, supernovae, and the collisions of stars.
Stellar surface abundances---the abundances measured by taking a spectrum of a stellar photosphere---are thought to deliver a relatively unprocessed record of the element abundances in the gas from which the star formed \citep[though see e.g.,][]{pinsonneault2001, oh2018, souto2019, vincenzo2021b}.
These birth abundances were set by a combination of nucleosynthetic processes involved in making heavy atomic nuclei, and astrophysical processes involved in delivering atoms from stellar interiors to star-formation sites \citep[e.g.,][]{johnsonja2020}.
Thus nuclear physics and a wide swath of astrophysics are critically intertwined in our understanding of stellar surface abundances, motivating theoretical, experimental, and observational work.

@@

\paragraph{3. Non-negativity}
All process vector components for all elements are non-negative and all process amplitudes are non-negative.
This assumption implies that the elements considered here are only produced, and not ever destroyed, by the $K$ processes (relative to hydrogen).
This makes the model similar to a non-negative matrix factorization (\citealt{nnmf, nnmf2}). In \name{}, this assumption is enforced by requiring that the process vector components and amplitudes are always greater than or equal to zero, such that
\begin{equation} \label{eq:constraint_zero}
    q_{k,j}^Z \geq 0 \; \forall \; Z,k,j \quad \text{and} \quad A_i^k \geq 0 \; \forall \; k,i ~.
\end{equation}

@@

Because we don't want to be too drawn or influenced by outlier points, we don't use the observed errors $\sigma_{{\rm obs},ij}$ in the likelihood, but instead we soften them in the spirit of iteratively reweighted least squares (\citealt{irls}):
\begin{equation}\label{eq:inflate_ivar}
    \frac{1}{\sigma_{ij}^2} = \frac{Q^2/\sigma_{{\rm obs},ij}^2}{Q^2 + (\xh_{ij} - m_{ij})^2 /\sigma_{{\rm obs},ij}^2 },    
\end{equation}
where $Q$ is a softening parameter, which is set to $Q=5$.

@@

\bibliography{kpm}{}
\bibliographystyle{aasjournal}
